---
title: "HW 5.2"
format: pdf
editor: visual
---

Setup:

```{r}
```

### Question 1

a). Given Y1 and Y2 that are both independent variables, the correlation should be 0. That is because the mean of Y1 and Y2 are separate from each other, so E(Y1, Y2) results in E(Y1) \* E(Y2). $Corr(Y_1, Y_2) = \frac{E(Y_1Y_2) - E(Y_1)E(Y_2)}{\sqrt{Var(Y_1)Var(Y_2)}} -> \frac{E(Y_1)E(Y_2) - E(Y_1)E(Y_2)}{\sqrt{Var(Y_1)Var(Y_2)}} -> 0$

b). The normal distribution has the unique property that if Y1 and Y2 are uncorrelated normal random variables, then they are independent. Usually correlation does not relate to independence because correlation only captures linear relationships, but non-linear relationships. However, in the unique case of a joint normal distribution this holds true due to the symmetric bell-surface of the joint normal distribution. It has no capability to embed any relationship other than linear ones, so if there is no correlation, then that must mean there are no relationships, and therefore independent.

### Question 2

a). The simplification of $Var(Y_1,Y_2)$ is (b) because when converting variance into $\sum^k_{j=1}\sum^k_{l=1}a_ja_lCov(Y_j, Y_l)$ you get $Cov(Y_1, Y_1) + Cov(Y_2, Y_2) + Cov(Y_1, Y_2) + Cov(Y_2, Y_1)$. Covariance of the same variable is just variance, and order of covariance input to the function does not matter. Therefore, we can simplify the equation to the answer (b), that is $Var(Y_1) + Var(Y_2) + 2Cov(Y_1, Y_2)$.

### Question 3

a). The formula would be $f(y;100,4) = \prod^{500}_{j=1}f(y_j;100,4) = \frac{1}{\sqrt{8\pi}}\exp(-\frac{1}{8} \sum^{500}_{j=1} (y_j-100)^2)$

```{r}
##########
# Part A #
##########
set.seed(123457) # For reproducability 
n <- 500
q3_mean <- 100
q3_var <- 4
samples <- rnorm(n, 100, 2)




```

b). The mean and variance are calculated below:

```{r}

# Because the mean for each sample is the same as that of
#the normal distribution the result would be as follows:
cat("The Summation Mean Is: ", n * q3_mean)

# The same logic can be applied for variance.
cat("\nThe Summation Variance Is: ", n * q3_var)


```

### Question 4

a). Given $Y_i, X_{i1}, X_{i2}$ for $n$ rows, the normal MLR model with intercept $\beta_0$ is as follows: $y = \beta_0 + \beta_1X_1 + \beta_2X_2$. To obtain each optimal beta, we must take partial derivatives of the optimization equation $S(\beta_0,\beta_1; Data) = \sum^n_{i=1}(y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2})^2$: $\newline \beta_{0,LSE} = \frac{\partial}{\partial \beta_0} = \sum^n_{i=1}(y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2})^2 \Rightarrow u = y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2}, u'=-1 \Rightarrow \newline -2\sum^n_{i=1}(y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2})$

$\beta_{1,LSE} = \frac{\partial}{\partial \beta_1} = \sum^n_{i=1}(y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2})^2 \Rightarrow u = y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2}, u'=-x_{i1} \Rightarrow \newline -2\sum^n_{i=1}(y_ix_{i1} - \beta_0x_{i1} - \beta_1x_{i1}^2 - \beta_2x_{i2}x_{i1})$

$\beta_{2,LSE} = \frac{\partial}{\partial \beta_2} = \sum^n_{i=1}(y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2})^2 \Rightarrow u = y_i - \beta_0 - \beta_1x_{i1} - \beta_2x_{i2}, u'=-x_{i2} \Rightarrow \newline -2\sum^n_{i=1}(y_ix_{i2} - \beta_0x_{i2} - \beta_1x_{i1}x_{i2} - \beta_2x_{i2}^2)$

### Question 5

```{r}

set.seed(123457)
n <- 80
beta_0 <- 2
beta_1 <- 0.5
beta_2 <- 0.85

error <- rnorm(n, mean = 0, sd = 0.5)
gen_x_1 <- rnorm(n, mean = 3, sd = 1)
gen_x_2 <- rbinom(n, size=1, prob=0.6)

y_val <- beta_0 + beta_1* gen_x_1 + beta_2*gen_x_2 + error

df = data.frame(x=gen_x_1,
                x2=gen_x_2,
                y=y_val)

lm(y ~ x + x2, data=df)



# y ~ beta_0 + beta_1 * x + beta_2 * x2 + error
#for (i in seq_along(gen_x_1)){
  # cat("i: ", i, "val: ", beta_0 + beta_1 * gen_x_1[i] + beta_2 * gen_x_2[i] + error[i], "\n")
  #y_val[i] <- beta_0 + beta_1* gen_x_1[i] + beta_2*gen_x_2[i] + error[i]
#}

```
